---
title: "IE 492 MCP First Look"
author: "Alican Yılmaz"
date: "3/29/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
 #setwd("/Users/Alican/Documents/Github/ie492-/Datasets")

```

## R Markdown

```{r READ, warning=FALSE, message=FALSE}
library('ggplot2')
library('forecast')
library('tseries')
library("data.table")
library("lubridate")
library("zoo")
library("plotly")
```


```{r}

mcp_raw = fread('MCP-01012017-29032021.csv', header=TRUE)
mcp_raw[,Date:=as.Date(Date,"%d/%m/%Y")]
```


```{r}
  daily_MCP=mcp_raw[,list(daily_MCP=mean(`PTF (USD/MWh)`,na.rm=T)),by=list(Date)]
```

```{r}
plt<-ggplot(daily_MCP, aes(Date, daily_MCP)) + geom_line()  + labs(x="Date",y="PTF (USD/MWh)",title="Daily Mean MCP between 2017-2021")

ggplotly(plt)
```

Variance not constant, there is trend for sure. Also, outliers exist. Transformation is necessary?

Currency also has an effect maybe? (2018 Currency crisis)

```{r acf/pacf}

ggAcf(daily_MCP$daily_MCP, lag=48,main="Daily MCP Time Series ACF")


```


Seasonality at weekly level.



```{r}
# Plotting a portion of daily data
ggplot(data = daily_MCP[1:28], aes(x = Date, y =daily_MCP )) +
  geom_line(color = "darkred") +
  labs(title = "Daily MCP(01.01.2017-28.01.2017) ",
       x = "Date",
       y= "MCP TL/(MWh)") +
  scale_x_date(date_minor_breaks = "1 day", date_breaks = "2 days", date_labels = "%d %b") +
  theme_minimal()
```




```{r}
plt<-ggplot(data = mcp_raw[1:96], aes(x = index(mcp_raw[1:96]), y = `MCP (TL/MWh)` )) +
  geom_line(color = "darkred") +
  labs(title = "Hourly MCP(01.01.2017-04.01.2017) ",
       x = "Date",
       y= "MCP TL/(MWh)") +
  theme_minimal()
ggplotly(plt)
```


There seems to be a seasonality at hourly level.

```{r ACF/PACF}
#market_clearing_price
ggAcf(mcp_raw$`MCP (TL/MWh)`, lag=72, main="Hourly MCP Time Series ACF")

```

Daily seasonality.

Weekends & holidays might affect.(But weekends might be redundant if weekly seasonality is put as input.)
Electricity Demand
Electricity Load
Temperature
Available capacity
Codes of Days
Supply & Demand of Electricity(Only shows one hour. How to download all?)
Weather conditions & Fuel Prices

```{r Data manipulation weekend and day code is added}
mcp_raw$day <- weekdays(as.Date(mcp_raw$Date))
mcp_raw$IsWeekend <- mcp_raw$day %in% c("Saturday","Sunday")
mcp_raw<-transform(mcp_raw ,IsWeekend = ifelse(mcp_raw$day %in% c("Saturday","Sunday") ,1,0 ))

```


# Web Scraping For Fuel Prices

```{r Fuel Prices in Turkey}
#install.packages("rvest")
#install.packages("dplyr")
library(rvest)
library(dplyr)
```


```{r}
link="https://www.tppd.com.tr/en/former-oil-prices?id=34&county=413&StartDate=01.01.2017&EndDate=29.03.2021"
page=read_html(link)

date<-page %>% html_nodes("td:nth-child(1)") %>% html_text()
fuel_price<-as.numeric(page %>% html_nodes("td:nth-child(6)") %>% html_text())
fuel_price_df<-data.frame(date,fuel_price,stringsAsFactors = FALSE)
fuel_price_df$date<-as.Date(fuel_price_df$date, format =  "%d %B %Y")
```


Some of the days are missing. We could interpolate them

```{r interpolating missing values of fuel prices}

ApproxFun <- approxfun(x = fuel_price_df$date, y = fuel_price_df$fuel_price)
Dates<-seq.Date(ymd("2017-01-01"),ymd("2021-04-09"),by=1)
LinearFit <- ApproxFun(Dates)


full_fuel_price_df<-data.frame(Date = Dates, fUEL_pRICE = LinearFit)

full_fuel_price_df<-full_fuel_price_df[full_fuel_price_df$Date<"2021-03-30",]

```




## ARIMA: Benchmark Purposes

Box-Jenkins Methodology will be used:

```{r}
adf.test(mcp_raw$`MCP (TL/MWh)`)
```

Data is stationary at 0.01 significance level.

```{r}
ggAcf(mcp_raw$`MCP (TL/MWh)`,lag.max = 168)
```

```{r}
ggPacf(mcp_raw$`MCP (TL/MWh)`,lag.max = 168)

```



```{r}
BoxCox.lambda(daily_MCP$`MCP (TL/MWh)`)
```

Lambda is close to 1, so no need transformation.






We know that there is weekly seasonality from the daily plot.

```{r}
daily_mcp_ts<-ts(daily_MCP$daily_MCP,frequency = 7)
fitted<-auto.arima(daily_mcp_ts,seasonal=TRUE,trace = TRUE)
checkresiduals(fitted)
```


```{r}
daily_mcp_ts<-ts(daily_MCP$daily_MCP,frequency = 7)
fitted<-arima(daily_mcp_ts,order=c(1,1,1),seasonal=c(2,0,0))
ggtsdisplay(fitted$residuals)
checkresiduals(fitted)
library(lmtest)
coeftest(fitted)
```

Model each hour separately for benchmark purpose.



## MAPE 

For one week:

Train: 2017-2020
Test: 3 Week

```{r}
# Error test
```

## DATA COLLECTING & RELATIONSHIP TESTING

HYPOTHESIS: ELECTRICITY DEMAND SIGNIFICANTLY AFFECTS MCP

### Getting Electricity Demand Data

```{r}

electricity_demand_df<-fread('RealTimeConsumption-01012017-12042021.csv', header=TRUE)
electricity_demand_df[,Date:=as.Date(Date,"%d.%m.%Y")]
electricity_demand_df[,`Consumption (MWh)`:=as.numeric(gsub(",", "", `Consumption (MWh)`))]
electricity_demand_df<-electricity_demand_df[Date<="2021-03-29",]

```



```{r creating data_frame includes all variables}
whole_Data<-mcp_raw
whole_Data$electricity_consumption<-electricity_demand_df$`Consumption (MWh)`
```

### Plotting Demand vs Price Relationship

```{r}
ggplot(whole_Data[Date<="2017-06-01",],aes(x=electricity_consumption, y =`MCP (TL/MWh)`))+
geom_point()+geom_smooth(method=lm)
```



Plotting at daily level, for each hour separately:

```{r}
ggplot(whole_Data,aes(x=electricity_consumption, y =`MCP (TL/MWh)`))+
geom_point()+geom_smooth(method=lm)+
  facet_wrap(vars(Hour))
```

### Regression Model for Electricity Demand



```{r}
lmMCP = lm(`MCP (TL/MWh)`~electricity_consumption, data = whole_Data) #Create the linear regression
summary(lmMCP) #Review the results
```

we can say that electricity consumption is statistically significant.However, R-square is quite low, meaning that variance can not be explained simply by consumption.

```{r}
ggtsdisplay(lmMCP$residuals)
```


This allows you to polynomial regression model of degree 3 of x


```{r}
#model <- lm(y ~ poly(x,3))
```


### Fuel_Prices vs MCP


```{r}
whole_Data_daily<-whole_Data %>%
  group_by(Date)
whole_Data_daily<-whole_Data_daily%>% summarise(
  mcp=mean(`MCP (TL/MWh)`))
whole_Data_daily$fuelprice<-full_fuel_price_df$fUEL_pRICE  

```



```{r}
whole_Data_daily%>%
  ggplot(.,aes(x=fuelprice,y=mcp))+
  geom_point()+geom_smooth(method=lm)
```

### Regression Model for Electricity Demand



```{r}
lmMCP = lm(mcp~fuelprice, data = whole_Data_daily) #Create the linear regression
summary(lmMCP) #Review the results
```

Statistically significant. But R^2 could be improved still.

## Hydro Production Analysis

Hypothesis: Real Time Hydro Production has an effect on MCP

### Getting Data


```{r}
production = fread('RealTimeGeneration-01012017-14032021.csv', header=TRUE)
production[,Date:=as.Date(Date,"%d.%m.%Y")]
production[,3:18] <- lapply(production[,3:18], function(x) as.numeric(gsub(",", "", x)))

```


```{r}
df<-whole_Data[Date<="2021-03-13",]
df$Dammed_Hydro_pr<-production[Date<="2021-03-13",`Dammed Hydro`]
```



```{r}
df%>%
  ggplot(.,aes(x=Dammed_Hydro_pr,y=`MCP (TL/MWh)`))+
  geom_point()+geom_smooth(method=lm)
```


```{r}
df%>%
  ggplot(.,aes(x=Dammed_Hydro_pr,y=`MCP (TL/MWh)`))+
  geom_point()+geom_smooth(method=lm) +facet_wrap(vars(Hour))
```


It is hard to tell the effect of dammed hydro production on MCP as it is hour dependent.

```{r}
lmMCP = lm(`MCP (TL/MWh)`~Dammed_Hydro_pr, data = df) #Create the linear regression
summary(lmMCP) #Review the results
```

R^2 is almost zero. 



## Bilateral Contracts

Hypothesis: Bilateral Contracts Total Bid Quantity has an effect on price.

```{r}
Bilateral_df = fread('BilateralContractsBid-01012017-29032021.csv', header=TRUE)
Bilateral_df[,Date:=as.Date(Date,"%d/%m/%Y")]
Bilateral_df[,3] <- lapply(Bilateral_df[,3], function(x) as.numeric(gsub(",", "", x)))
```

```{r}
whole_Data$'bilateral(MWh)'<-Bilateral_df[,3]


whole_Data[Date>="2020-01-01",]%>%
  ggplot(.,aes(x=`bilateral(MWh)`,y=`MCP (TL/MWh)`))+
  geom_point()+geom_smooth(method=lm)


```




```{r}
lmMCP = lm(`MCP (TL/MWh)`~`bilateral(MWh)`, data = whole_Data[Date>="2020-01-01",]) #Create the linear regression
summary(lmMCP) #Review the results
```

# Corrplot

```{r}
library(corrplot)

M<-cor(whole_Data[,c(3,8)])

corrplot(M, method="number")
```

## Installed Capacity vs MCP

```{r}
capacity=fread("EAK-01012017-29032021.csv",header=TRUE)
capacity[,Date:=as.Date(Date,"%d.%m.%Y")]

capacity[,`Total (MWh)`:=as.numeric(gsub(",", "",`Total (MWh)`))]
head(capacity)
  
```


```{r}
whole_Data$InstalledCapacity<-capacity$`Total (MWh)`
head(whole_Data)
```

```{r}
whole_Data%>%
  ggplot(.,aes(x=InstalledCapacity,y=`MCP (TL/MWh)`))+
  geom_point()+geom_smooth(method=lm) #+facet_wrap(vars(Hour))
```


```{r}
whole_Data %>%
  group_by(Date)%>%
  summarise(
    daily_MCP = sum(`MCP (TL/MWh)`),
  daily_capacity = sum(InstalledCapacity))%>%
  ggplot(.,aes(x=daily_capacity,y=daily_MCP))+
  geom_point()+geom_smooth(method=lm)
```






## Agraggate Daily Data





## Daily Mean Electricity Consumption Data


```{r}
daily_whole_data<-whole_Data %>% 
  group_by(Date) %>% 
  summarise(across(where(is.numeric), list(mean = mean, sum = sum)))
```

```{r}
model <- lm(`MCP (TL/MWh)_sum` ~ poly(InstalledCapacity_sum,3),data=daily_whole_data)
summary(model)
plot(fitted(model),residuals(model))
```
## Outlier Analysis



```{r}
daily_whole_data%>%
  ggplot(.,aes(x=Date,y=`MCP (TL/MWh)_mean`))+
  geom_line(color = "darkred") +
  labs(title = "Daily MCP",
       x = "Date",
       y= "MCP TL/(MWh)") +
  theme_minimal()
```

## Daily Hydro Production

```{r}
df_aggregated<-df%>% 
  group_by(Date) %>% 
  summarise(across(where(is.numeric), list(mean = mean, sum = sum)))
df_aggregated
```

```{r}
df_aggregated%>%ggplot(.,aes(x=`Dammed_Hydro_pr_sum`,y=`MCP (TL/MWh)_sum`))+
  geom_point()+geom_smooth(method=lm)
```



```{r}
`MCP (USD/MWh)`<-ts(df_aggregated$`PTF (USD/MWh)_mean`,start=c(2017,1), frequency=365)
`Dammed_Hydro_Pr(Mwh)`<-ts(ts(df_aggregated$Dammed_Hydro_pr_mean),start=c(2017,1), frequency=365)

autoplot(cbind(`MCP (USD/MWh)`,`Dammed_Hydro_Pr(Mwh)`),facets=TRUE)+
  labs(x = "Date", y ="",title = "Market Clearing Price vs Dammed Hydro Production Level 2017-2021")+
  scale_x_yearmon(n=5)

```


```{r}
library(corrplot)

M<-cor(df_aggregated[,c(2,10)])

corrplot(M, method="number")
```



```{r}
df_hydro<-df_aggregated[,c("MCP (TL/MWh)_mean","Dammed_Hydro_pr_mean","MCP (TL/MWh)_sum","Dammed_Hydro_pr_sum")]


```


```{r}
Q <- quantile(df_hydro$Dammed_Hydro_pr_mean, probs=c(.25, .75), na.rm = FALSE)
Q
```


```{r}
eliminated<- subset(df_hydro, df_hydro$Dammed_Hydro_pr_mean < (Q[1]) | df_hydro$Dammed_Hydro_pr_mean > (Q[2]))
eliminated
```

```{r}
#eliminated %>% ggplot(.,aes(x=`Dammed_Hydro_pr_mean`,y=`MCP (TL/MWh)_mean`))+
#  geom_point()+geom_smooth(method=lm)
```



## Planned Production vs Planned Load

```{r}

planned_prod = fread('Final+Daily+Production+Program-01012017-29032021.csv', header=TRUE)
planned_prod[,Date:=as.Date(Date,"%d.%m.%Y")]
planned_prod[,c("Date","Hour",`Total (MWh)`)]
planned_load=fread('LoadForecast-01012017-29032021.csv', header=TRUE)
planned_load[,Date:=as.Date(Date,"%d/%m/%Y")]
new_df_prod_load<-merge.data.frame(planned_load,planned_prod,by=c("Date","Hour"))
new_df_prod_load<-data.table(new_df_prod_load)
new_df_prod_load[,`Load Forecast (MWh)`:=as.numeric(gsub(",", "", `Load Forecast (MWh)`))]
new_df_prod_load[,`Total (MWh)`:=as.numeric(gsub(",", "", `Total (MWh)`))]
new_df_prod_load$ratio<-new_df_prod_load$`Load Forecast (MWh)`/new_df_prod_load$`Total (MWh)`
```


```{r}

df_comb<-merge(new_df_prod_load[,c(1,2,17)],mcp_raw[,c(1:3)],by=c("Date","Hour"))

df_comb %>% ggplot(.,aes(x=`MCP (TL/MWh)`,y=ratio))+
  geom_point()
```

```{r}
library(tidyverse)
df_comb %>%
  mutate(pct_change = (lead(ratio)/ratio - 1) * 100)%>% ggplot(.,aes(x=`MCP (TL/MWh)`,y=pct_change))+
  geom_point()
```

# Wind ratio & Hydro ratio analysis

```{r}
mcp_raw<-mcp_raw[Date<="2021-03-13",]
production<-production[Date<="2021-03-13",c("Date","Hour","Total (MWh)","Dammed Hydro","Wind")]
merged_data<-merge(mcp_raw,production,by =c("Date","Hour"))
merged_data<-merged_data %>%
  mutate(pct_wind = (Wind/`Total (MWh)`) * 100, pct_dammed_hydro=(`Dammed Hydro`/`Total (MWh)`) * 100)
```


```{r}
merged_data2<-merged_data[Date>="2021-02-20",]

autoplot(cbind(ts(merged_data2$`MCP (TL/MWh)`),ts(merged_data2$pct_dammed_hydro)),facets=TRUE)
```

```{r}
library(corrplot)

M<-cor(merged_data[,c(3,9,12)])

corrplot(M, method="number")
```




### Differencing

```{r}

merged_data<-merged_data%>%
  mutate(diff_perc_hydro=c(NA,diff(pct_dammed_hydro)),diff_perc_wind=c(NA,diff(pct_dammed_hydro)),diff_MCP=c(NA,diff(`MCP (TL/MWh)`)))

cor(merged_data$diff_MCP,merged_data$diff_perc_hydro, use="complete.obs")
```


```{r}
merged_data3<-merged_data[Date>="2021-02-20",]

autoplot(cbind(ts(merged_data3$diff_MCP),ts(merged_data3$diff_perc_wind),ts(merged_data3$diff_perc_hydro)),facets=TRUE)
```

## Percentage of Renewable

## Granger test of causality
 First, be sure that both are stationary. Order(lag) is important!
 
```{r}
library(lmtest)
grangertest(merged_data$`MCP (TL/MWh)`~merged_data$pct_dammed_hydro, order=1)
```

```{r}
#install.packages("vars")
library(vars)
causal_df<-cbind(merged_data$pct_dammed_hydro,merged_data$`MCP (TL/MWh)`)
causal_car<-VAR(causal_df,type="const", lag.max=10 , ic = "AIC")
causality(causal_car,cause="y1")$Granger
```

## Wind percentage effect

```{r}
merged_data3<-merged_data[Date>="2021-02-20",]

autoplot(cbind(ts(merged_data3$`MCP (TL/MWh)`),ts(merged_data3$pct_wind)),facets=TRUE)
```


```{r}
merged_data4<-merged_data[Date<="2021-01-20" & Date>="2020-01-16",]

autoplot(cbind(ts(merged_data4$`MCP (TL/MWh)`),ts(merged_data4$pct_wind),ts(merged_data4$pct_dammed_hydro)),facets=TRUE)
```

```{r}
library(vars)
causal_df<-cbind(merged_data$pct_wind,merged_data$`MCP (TL/MWh)`)
causal_car<-VAR(causal_df,type="const", lag.max=10 , ic = "AIC")
causality(causal_car,cause="y2")$Granger
```


```{r}
merged_all = fread('merged.csv', header=TRUE)
merged_all[,Date:=as.Date(Date,"%d-%m-%Y")]
```

# NATURAL GAS PROPORTION

```{r}
merged_data4<-merged_all[Date<="2017-06-20" & Date>="2017-06-15",] 

autoplot(cbind(ts(merged_data4$MCP..TL.MWh.),ts(merged_data4$GProportion.NatGas),ts(merged_data4$GProportion.Wind),ts(merged_data4$GProportion.DammedH),ts(merged_data4$GProportion.Lignite),ts(merged_data4$GProportion.ImportC)),facets=TRUE)
```

## Renewable proportion analysis

```{r}
merged_all<-merged_all%>%
  mutate(pct_renewable = (GProportion.DammedH+GProportion.Biomass+GProportion.Geot+GProportion.River+GProportion.Solar))
```


```{r}
merged_data4<-merged_all[Date<="2018-10-20" & Date>="2018-05-15",] 
merged_data4<-merged_data4[Hour==12]
`MCP (USD/MWh)`<-ts(merged_data4$PTF..USD.MWh.,start=c(2018,5), frequency=365)
Pct_Renewable<-ts(merged_data4$pct_renewable,start=c(2018,5),frequency = 365)
autoplot(cbind(`MCP (USD/MWh)`,pct_renewable),facets=TRUE)+
   labs(x = "Date", y ="",title = "Market Clearing Price vs Renewable to Production Ratio in 2018")+
  scale_x_yearmon(n=12)


#renewable ratio outlier
merged_data4<-merged_all
merged_data4<-merged_data4[Hour==12]


Q <- quantile(merged_data4$pct_renewable, probs=c(.05, .95), na.rm = TRUE)
Q

merged_data4[,pct_renewable_small:= as.numeric(merged_data4$pct_renewable < (Q[1]))]
merged_data4[,pct_renewable_great:= as.numeric(merged_data4$pct_renewable > (Q[2]))]

plt<-ggplot(merged_data4,aes(x=as.factor(pct_renewable_small),y=PTF..USD.MWh.)) + geom_boxplot(length= 0.5) 
ggplotly(plt)

plt<-ggplot(merged_data4,aes(x=as.factor(pct_renewable_great),y=PTF..USD.MWh.)) + geom_boxplot(length= 0.5) 
ggplotly(plt)

```



```{r}
cor(merged_all$MCP..TL.MWh.,merged_all$pct_renewable,use = "pairwise.complete.obs", method = c("pearson"))
```

```{r}
M<-cor(merged_all[,c("MCP..TL.MWh.","pct_renewable")],use ="pairwise.complete.obs",method="pearson" )

corrplot(M, method="number")
```


## Special Days

Weekend,day codes, special days.

```{r}
mcp_raw
```

```{r}
model<-lm(formula = `PTF (USD/MWh)`~ -1 + as.factor(day), 
data = mcp_raw)
summary(model)
checkresiduals(model)
```

```{r}
mcp_raw[, lag_24:= shift(`PTF (USD/MWh)`,24)]
mcp_raw[, trend:= seq(1:nrow(mcp_raw))]
model<-lm(formula = `PTF (USD/MWh)`~ -1 + as.factor(day)+lag_24+trend, 
data = mcp_raw)
summary(model)
checkresiduals(model)
```
## outliers

```{r}
length(model$residuals)

```





```{r}
mcp_raw[,residuals := NA]
mcp_raw$residuals[25:nrow(mcp_raw)]<-model$residuals
Q <- quantile(mcp_raw$residuals, probs=c(.05, .95), na.rm = TRUE)
Q

outliers<- subset(mcp_raw, mcp_raw$residuals < (Q[1]) | mcp_raw$residuals > (Q[2]))
outliers
mcp_raw[,outlier_small:= as.numeric(mcp_raw$residuals < (Q[1]))]
mcp_raw[,outlier_great:= as.numeric(mcp_raw$residuals > (Q[2]))]
```

```{r}
outliers<- subset(model$residuals, model$residuals < (Q[1]) | model$residuals > (Q[2]))
#outliers
```

```{r}
model<-lm(formula = `PTF (USD/MWh)`~ -1 + as.factor(day)+lag_24+trend+outlier_small+outlier_great, 
data = mcp_raw)
summary(model)
checkresiduals(model)
```

## add special days to the dataset as dummy 

```{r}
mcp_raw[,is_yilbasi := 0]
mcp_raw[mcp_raw$Date %in% yilbasi,is_yilbasi:=1]

mcp_raw[,is_sevgililergunu := 0]
mcp_raw[mcp_raw$Date %in% sevgililergunu,is_sevgililergunu:=1]



mcp_raw[,is_cocukbayrami := 0]
mcp_raw[mcp_raw$Date %in% cocukbayrami,is_cocukbayrami:=1]

mcp_raw[,is_iscibayrami := 0]
mcp_raw[mcp_raw$Date %in% iscibayrami,is_iscibayrami:=1]

mcp_raw[,is_annelergunu := 0]
mcp_raw[mcp_raw$Date %in% annelergunu,is_annelergunu:=1]

mcp_raw[,is_genclikbayrami := 0]
mcp_raw[mcp_raw$Date %in% genclikbayrami,is_genclikbayrami:=1]

mcp_raw[,is_babalargunu := 0]
mcp_raw[mcp_raw$Date %in% babalargunu,is_babalargunu:=1]

mcp_raw[,is_zaferbayrami := 0]
mcp_raw[mcp_raw$Date %in% zaferbayrami,is_zaferbayrami:=1]

mcp_raw[,is_cumhuriyet := 0]
mcp_raw[mcp_raw$Date %in% cumhuriyet,is_cumhuriyet:=1]

mcp_raw[,is_kadinlargunu := 0]
mcp_raw[mcp_raw$Date %in% kadinlargunu,is_kadinlargunu:=1]


mcp_raw[,is_ramazanbayrami := 0]
mcp_raw[mcp_raw$Date %in% ramazanbayrami,is_ramazanbayrami:=1]

mcp_raw[,is_kurbanbayrami := 0]
mcp_raw[mcp_raw$Date %in% kurbanbayrami,is_kurbanbayrami:=1]

mcp_raw[,is_uzunbayram := 0]
mcp_raw[mcp_raw$Date %in% uzunbayram,is_kurbanbayrami:=1]

mcp_raw[,is_uzunbayram := 0]
mcp_raw[mcp_raw$Date %in% uzunbayram,is_kurbanbayrami:=1]

mcp_raw[,is_uzunbayram := 0]
mcp_raw[mcp_raw$Date %in% uzunbayram,is_kurbanbayrami:=1]

mcp_raw[,is_okulacilisi := 0]
mcp_raw[mcp_raw$Date %in% okulacilisi,is_okulacilisi:=1]
okulkapanisi
mcp_raw[,is_okulkapanisi := 0]
mcp_raw[mcp_raw$Date %in% okulkapanisi,is_okulkapanisi:=1]

mcp_raw[,is_semester := 0]
mcp_raw[mcp_raw$Date %in% semester,is_semester:=1]

mcp_raw[,is_kandil := 0]
mcp_raw[mcp_raw$Date %in% kandil,is_kandil:=1]

ramazanbayramilkgun
mcp_raw[,is_ramazanbayramilkgun := 0]
mcp_raw[mcp_raw$Date %in% ramazanbayramilkgun,is_ramazanbayramilkgun:=1]
ramazanbayramisonrasi
mcp_raw[,is_ramazanbayramisonrasi := 0]
mcp_raw[mcp_raw$Date %in% ramazanbayramisonrasi,is_ramazanbayramisonrasi:=1]

mcp_raw[,is_kurbanbayramilkgun := 0]
mcp_raw[mcp_raw$Date %in% kurbanbayramilkgun,is_kurbanbayramilkgun:=1]

mcp_raw[,is_ramazangunler := 0]
mcp_raw[mcp_raw$Date %in% ramazangunler,is_ramazangunler:=1]

mcp_raw[,is_yilbasioncesi := 0]
mcp_raw[mcp_raw$Date %in% yilbasioncesi,is_yilbasioncesi:=1]

mcp_raw[,is_kurbanoncesigunler := 0]
mcp_raw[mcp_raw$Date %in% kurbanoncesigunler,is_kurbanoncesigunler:=1]

mcp_raw[,is_ramazanilkgun := 0]
mcp_raw[mcp_raw$Date %in% ramazanilkgun,is_ramazanilkgun:=1]

mcp_raw[,is_ramazanoncesigunler := 0]
mcp_raw[mcp_raw$Date %in% ramazanoncesigunler,is_ramazanoncesigunler:=1]

mcp_raw[,is_arife := 0]
mcp_raw[mcp_raw$Date %in% arife,is_arife:=1]

mcp_raw[,is_onbestemmuz := 0]
mcp_raw[mcp_raw$Date %in% onbestemmuz,is_onbestemmuz:=1]

mcp_raw[,is_holiday := 0]
mcp_raw[mcp_raw$Date %in% holiday,is_holiday:=1]

mcp_raw[,is_outlierdays := 0]
mcp_raw[mcp_raw$Date %in% outlierdays,is_outlierdays:=1]

mcp_raw[,is_elections:= 0]
mcp_raw[mcp_raw$Date %in% elections,is_elections:=1]

mcp_raw[,is_examination:= 0]
mcp_raw[mcp_raw$Date %in% examination,is_examination:=1]

mcp_raw[,is_curfew:= 0]
mcp_raw[mcp_raw$Date %in% curfew,is_curfew:=1]

mcp_raw[,is_single_day_holidays:= 0]
mcp_raw[mcp_raw$Date %in% single_day_holidays,is_single_day_holidays:=1]

mcp_raw[,is_school_start:= 0]
mcp_raw[mcp_raw$Date %in% school_start,is_school_start:=1]

mcp_raw[,is_school_end:= 0]
mcp_raw[mcp_raw$Date %in% school_end,is_school_end:=1]

mcp_raw[,is_school_period:= 0]
mcp_raw[mcp_raw$Date %in% school_period,is_school_period:=1]

mcp_raw[,is_school_days:= 0]
mcp_raw[mcp_raw$Date %in% school_days,is_school_days:=1]

mcp_raw[,is_semester_days:= 0]
mcp_raw[mcp_raw$Date %in% semester_days,is_semester_days:=1]

mcp_raw[,is_winter_break:= 0]
mcp_raw[mcp_raw$Date %in% winter_break,is_winter_break:=1]

mcp_raw[,is_summer_break:= 0]
mcp_raw[mcp_raw$Date %in% summer_break,is_summer_break:=1]

install.packages("writexl")

library("writexl")
write_xlsx(mcp_raw,"/Users/alicanyilmaz/Desktop///mcp_special_days.xlsx")

```


```{r}
# +outlier_small+outlier_great is deleted
model<-lm(formula = `PTF (USD/MWh)`~ ., 
data = mcp_raw[,-c("Date","Hour","MCP (TL/MWh)","PTF (EUR/MWh)","residuals","outlier_small"            ,"outlier_great","is_holiday","is_single_day_holidays","generation/consumption","outlier_ratio_small","outlier_ratio_great")])
summary(model)
checkresiduals(model)
```

# Special Days Price Spike Relation


```{r}

nrow(mcp_raw[is_ramazanbayrami==1])
nrow(mcp_raw[is_ramazanbayrami==1 & outlier_small==1])

```








## h0= production/consumption has an effect on mcp


```{r}
merged = fread('merged.csv', header=TRUE)
merged2<-merged[,c("Date","Hour","PTF..USD.MWh.","Generation/Consumption")]
# bi makalede oran belli bi değerdeyse (0.9 Civarı) iyi falan yazıyordu. Hangisiydi hatırlayamadım.Şimdilik bunu gözardı edelim.



corrplot(cor(merged2[,c("PTF..USD.MWh.","Generation/Consumption")],use = "pairwise.complete.obs"), method="number")


```


```{r}
merged2<-merged2[complete.cases(merged2),]
Q <- quantile(merged2$PTF..USD.MWh., probs=c(.05, .95), na.rm = TRUE)
Q

merged2[,outlier_ptf_small:= as.numeric(merged2$PTF..USD.MWh. < (Q[1]))]
merged2[,outlier_ptf_great:= as.numeric(merged2$PTF..USD.MWh. > (Q[2]))]

Q <- quantile(merged2$`Generation/Consumption`, probs=c(.05, .95), na.rm = TRUE)
Q

merged2[,outlier_ratio_small:= as.numeric(merged2$`Generation/Consumption` < (Q[1]))]
merged2[,outlier_ratio_great:= as.numeric(merged2$`Generation/Consumption` > (Q[2]))]

corrplot(cor(merged2[,c("outlier_ptf_small","outlier_ptf_great","outlier_ratio_small","outlier_ratio_great")],use = "pairwise.complete.obs"), method="number")


res.aov2 <- aov( PTF..USD.MWh.~ outlier_ratio_small +outlier_ratio_great, data = merged2)
summary(res.aov2)

```


```{r}
plt<-ggplot(merged2[Hour==21],aes(x=as.factor(outlier_ratio_small),y=PTF..USD.MWh.)) + geom_boxplot(length= 0.5) 
ggplotly(plt)
```

```{r}
plt<-ggplot(merged2[Hour==16],aes(x=as.factor(outlier_ratio_great),y=PTF..USD.MWh.)) + geom_boxplot(length= 0.5) 
ggplotly(plt)
```

#Residuallar üzerinden gitmek daha mantıklı olabilir. Çünkü ratio büyük de olsa sabah saatlerinde ptf küçük olacak.Ya da saat saat bak
```{r}
library(dplyr)
merged2<-merged2[1:36792,]
mcp_raw$'generation/consumption'<-merged2$`Generation/Consumption`
mcp_raw$outlier_ratio_small<-merged2$outlier_ratio_small
mcp_raw$outlier_ratio_great<-merged2$outlier_ratio_great

plt<-ggplot(mcp_raw[],aes(x=as.factor(outlier_ratio_small),y=residuals)) + geom_boxplot(length= 0.5) 
ggplotly(plt)
```

## Installed capacity/planned laod

```{r}

#new_df<-join(merged[Date<="2021-03-13",c("Date","Hour","PTF..USD.MWh.","EAK.Total (MWh)")],mcp_raw,by=c("Date","Hour"))
```


## planned load/planned consumption ratio 

```{r}

new_df_prod_load<-new_df_prod_load[Date<="2021-03-13",]
mcp_raw$load_consumption_ratio<-new_df_prod_load$ratio

mcp_raw[,is_l_c_ratio_big := 0]
mcp_raw[load_consumption_ratio>=1,is_l_c_ratio_big:=1.1]

plt<-ggplot(mcp_raw[Hour=="12:00"],aes(x=as.factor(is_l_c_ratio_big),y=`PTF (USD/MWh)`)) + geom_boxplot(length= 0.5) 
ggplotly(plt)


```





#ARIMA BENCHMARK ANALYSIS


```{r}
mcp_raw = fread('MCP-01012017-29032021.csv', header=TRUE)
mcp_raw[,Date:=as.Date(Date,"%d/%m/%Y")]
```

```{r}
mcp_raw_train<-mcp_raw[Date<"2019-01-01"]
mcp_raw_test<-mcp_raw[Date>="2019-01-01" & Date<="2020-01-01"]
mcp_raw_hourly<-mcp_raw_train[Hour=="12:00"] #choose an hour to be fitted
mcp_raw_ts<-ts(mcp_raw_hourly$`PTF (USD/MWh)`,frequency = 7)
fitted<-auto.arima(mcp_raw_ts,seasonal=TRUE,trace = TRUE)
checkresiduals(fitted)
```
```{r}

fitted<-arima(mcp_raw_ts,order=c(1,1,1),seasonal=c(2,0,0))
ggtsdisplay(fitted$residuals)
checkresiduals(fitted)
library(lmtest)
coeftest(fitted)
```

```{r}
y<-mcp_raw_test[Hour=="12:00"]$`PTF (USD/MWh)`

refit<-Arima(y,model=fitted)
ACC<-accuracy(refit)


```

```{r}
print(ACC[,"RMSE"])
print(ACC[,"MAE"])
```











### ARIMA ANALYSIS

```{r}

mcp_raw = fread('MCP-01012017-29032021.csv', header=TRUE)
mcp_raw[,Date:=as.Date(Date,"%d/%m/%Y")]
mcp_raw_test<-mcp_raw[Date>="2019-01-01" & Date<="2020-01-01"]
```

# define the function which predicts on a monthly basis
```{r}
fitted
```

```{r}
arima_monthly_slide <- function(s_Hour) {
  #january  
  mcp_raw_train<-mcp_raw[Date<"2019-01-01"]
  
  mcp_raw_hourly<-mcp_raw_train[Hour==s_Hour] #choose an hour to be fitted
  mcp_raw_ts<-ts(mcp_raw_hourly$`PTF (USD/MWh)`,frequency = 7)
  
  fitted<-auto.arima(mcp_raw_ts,seasonal=TRUE,trace = FALSE)
  fitted
  mcp_raw_test_<-mcp_raw_test[Hour==s_Hour & Date <"2019-02-01"] # 1 month data to be predicted
  mcp_raw_ts<-ts(mcp_raw_test_$`PTF (USD/MWh)`,frequency = 7)
  refit<- Arima(mcp_raw_ts, model=fitted)
  
  
  
}
mcp_raw_test<-mcp_raw_test[Date<"2020-01-01"]

mcp_raw_test<-arima_monthly_slide("00:00")
mcp_raw_test<-arima_monthly_slide("01:00")
mcp_raw_test<-arima_monthly_slide("02:00")
mcp_raw_test<-arima_monthly_slide("03:00")
mcp_raw_test<-arima_monthly_slide("04:00")
mcp_raw_test<-arima_monthly_slide("05:00")
mcp_raw_test<-arima_monthly_slide("06:00")
mcp_raw_test<-arima_monthly_slide("07:00")
mcp_raw_test<-arima_monthly_slide("08:00")
mcp_raw_test<-arima_monthly_slide("09:00")
mcp_raw_test<-arima_monthly_slide("10:00")
mcp_raw_test<-arima_monthly_slide("11:00")
mcp_raw_test<-arima_monthly_slide("12:00")
mcp_raw_test<-arima_monthly_slide("13:00")
mcp_raw_test<-arima_monthly_slide("14:00")
mcp_raw_test<-arima_monthly_slide("15:00")
mcp_raw_test<-arima_monthly_slide("16:00")
mcp_raw_test<-arima_monthly_slide("17:00")
mcp_raw_test<-arima_monthly_slide("18:00")
mcp_raw_test<-arima_monthly_slide("19:00")
mcp_raw_test<-arima_monthly_slide("20:00")
mcp_raw_test<-arima_monthly_slide("21:00")
mcp_raw_test<-arima_monthly_slide("22:00")
mcp_raw_test<-arima_monthly_slide("23:00")

```

# accuracy test 

```{r}
error_test <- function(actual, forecasted){
  n=length(actual)
  error = actual-forecasted
  mean=mean(actual)
  sd=sd(actual)
  bias = sum(error)/sum(actual)
  mape = sum(abs(error/actual))/n
  mad = sum(abs(error))#mae
  wmape = mad/mean
  rmse= sqrt(sum(error^2)/n) #buna g?re yap?caz
  mae=sum(abs(error))/n #
  MPE = sum(error/actual)/n
  df = data.frame(n,mean,sd,bias,mape,mad,wmape,rmse,mae,MPE)
  return(df)
}
```

## Data visualization 

### Hourly Analysis

```{r}
tail(mcp_raw_test)
```
ts to numeric conversion

```{r}
mcp_raw_test$predicted<-as.numeric(mcp_raw_test$predicted)
str(mcp_raw_test)
```

#hourly analysis

```{r}
mae_vec<-c(1:24)
rmse_vec<-c(1:24)
i=1
Hour_list<-c("00:00","01:00","02:00","03:00","04:00","05:00","06:00","07:00","08:00","09:00","10:00","11:00","12:00","13:00","14:00","15:00","16:00","17:00","18:00","19:00","20:00","21:00","22:00","23:00")
for (k in Hour_list){
  acc_data<-mcp_raw_test[Hour==k]
  x<-error_test(acc_data$`PTF (USD/MWh)`,acc_data$predicted)
  mae_vec[i]<-x$mae
  rmse_vec[i]<-x$rmse

  i=i+1
}
```


```{r}
mae_vec
rmse_vec
```








```{r}
wmape_vec<-c(1:24)
mad_vec<-c(1:24)
i=1
Hour_list<-c("00:00","01:00","02:00","03:00","04:00","05:00","06:00","07:00","08:00","09:00","10:00","11:00","12:00","13:00","14:00","15:00","16:00","17:00","18:00","19:00","20:00","21:00","22:00","23:00")
for (k in Hour_list){
  acc_data<-mcp_raw_test[Hour==k]
  acc_data$predicted<-as.numeric(acc_data$predicted)
  x<-error_test(acc_data$`PTF (USD/MWh)`,acc_data$predicted)
  wmape_vec[i]<-x$wmape
  mad_vec[i]<-x$mad

  i=i+1
}

```






```{r}

df_hourly_acc <- data.frame(Hour=Hour_list,
                 wmape=wmape_vec,
                 mad=mad_vec)

```

# Hourly Errors

```{r}
ggplot<-df_hourly_acc  %>% ggplot(aes(x=Hour, y=wmape, fill=Hour)) + geom_col() +theme_minimal() + theme(legend.position = "None",axis.text.x = element_text(angle = 45)) +labs(x="Hour",y="wmape by hour",title="WMAPE values of ARIMA model in 2019")

ggplotly(ggplot)
```

```{r}
ggplot<-df_hourly_acc  %>% ggplot(aes(x=Hour, y=mad, fill=Hour)) + geom_col() +theme_minimal() + theme(legend.position = "None",axis.text.x = element_text(angle = 45)) +labs(x="Hour",y="mad by hour",title="MAD values of ARIMA model in 2019")

ggplotly(ggplot)
```

## Seasonal Analysis


```{r}
mcp_raw_test
metseasons <- c(
  "01" = "Winter", "02" = "Winter",
  "03" = "Spring", "04" = "Spring", "05" = "Spring",
  "06" = "Summer", "07" = "Summer", "08" = "Summer",
  "09" = "Fall", "10" = "Fall", "11" = "Fall",
  "12" = "Winter"
)
metseasons[format(mcp_raw_test$Date, "%m")]
mcp_raw_test2<-mcp_raw_test[,season:=metseasons[format(mcp_raw_test$Date, "%m")]]
```

```{r}
#initially null vectors
seasonal_mae<-c(1:4)
seasonal_rmse<-c(1:4)

#update it for each season
acc_data<-mcp_raw_test2[season=="Fall"]
acc_data$predicted<-as.numeric(acc_data$predicted)
x<-error_test(acc_data$`PTF (USD/MWh)`,acc_data$predicted)
seasonal_mae[3]<-x$mae
seasonal_rmse[3]<-x$rmse


df_seasonal_acc <- data.frame(Season=c("Spring","Summer","Fall","Winter"),
                 mae=seasonal_mae,
                 rmse=seasonal_rmse)
df_seasonal_acc
mae_arima_Seasonal<-c(7.935285,5.052711,3.042565	,5.951492)
rmse_arima_Seasonal<-c(11.148063,9.238145,5.226218	,9.122256	)
```

```{r}
ggplot<-df_seasonal_acc  %>% ggplot(aes(x=Season, y=wmape, fill=Season)) + geom_col() +theme_minimal() + theme(legend.position = "None",axis.text.x = element_text(angle = 45)) +labs(x="Season",y="wmape",title="WMAPE values of ARIMA model in 2019")

ggplotly(ggplot)
```

```{r}
ggplot<-df_seasonal_acc  %>% ggplot(aes(x=Season, y=mad, fill=Season)) + geom_col() +theme_minimal() + theme(legend.position = "None",axis.text.x = element_text(angle = 45)) +labs(x="Season",y="mad",title="MAD values of ARIMA model in 2019")

ggplotly(ggplot)
```


# Special day analysis


```{r}
mcp_raw_test

mcp_raw_2019<-mcp_raw[Date>="2019-01-01" & Date<"2020-01-01",]

x_merged<-merge(mcp_raw_test,mcp_raw_2019)

#initially null vectors
specialday_wmape<-c(1:40)
specialday_mad<-c(1:40)

#update it for each season
special_days=c("is_yilbasi","is_sevgililergunu","is_cocukbayrami","is_iscibayrami"   ,    "is_annelergunu","is_genclikbayrami", "is_babalargunu", "is_zaferbayrami"  ,        "is_cumhuriyet"       ,     "is_kadinlargunu" ,         "is_ramazanbayrami"       
,"is_kurbanbayrami"   ,      "is_uzunbayram"       ,     "is_okulacilisi"  ,         "is_okulkapanisi"      ,   
"is_semester"     ,         "is_kandil"          ,      "is_ramazanbayramilkgun"  , "is_ramazanbayramisonrasi",
"is_kurbanbayramilkgun"   , "is_ramazangunler"     ,    "is_yilbasioncesi"      ,   "is_kurbanoncesigunler"   ,
 "is_ramazanilkgun"   ,      "is_ramazanoncesigunler" ,  "is_arife"             ,    "is_onbestemmuz"   ,      
"is_holiday"       ,        "is_outlierdays"      ,     "is_elections"      ,       "is_examination"   ,       
 "is_curfew",                "is_single_day_holidays" ,  "is_school_start"      ,    "is_school_end"    ,       
"is_school_period"   ,      "is_school_days"      ,     "is_semester_days"     ,    "is_winter_break"    ,     
"is_summer_break")
special_days_df<-subset(x_merged,select=-c(Date,Hour,`MCP (TL/MWh)`, `PTF (USD/MWh)`,`PTF (EUR/MWh)`, predicted,season))

########
k=1
for(i in colnames(special_days_df)){
  #gsub
  acc_data<-x_merged[i==1]
  acc_data$predicted<-as.numeric(acc_data$predicted)
  x<-error_test(acc_data$`PTF (USD/MWh)`,acc_data$predicted)
  specialday_wmape[k]<-x$wmape
  specialday_mad[k]<-x$mad
  k=k+1 
}
specialday_wmape

acc_data<-x_merged[is_semester_days==1]
acc_data$predicted<-as.numeric(acc_data$predicted)
x<-error_test(acc_data$`PTF (USD/MWh)`,acc_data$predicted)
specialday_wmape[38]<-x$wmape
specialday_mad[38]<-x$mad





specialday_wmape
specialday_mad

df_seasonal_acc <- data.frame(special_day=special_days,
                 wmape=specialday_wmape,
                 mad=specialday_mad)



```


```{r}
ggplot<-ggplot(df_seasonal_acc) +
    geom_point(aes(x = wmape, y = special_day))
ggplotly(ggplot)
```

```{r}
mcp_raw_test$day <- weekdays(as.Date(mcp_raw_test$Date)) #add week code column to data set
```

```{r}
#calculate mae/rmse of each day code
mae_vec<-c(1:7)
rmse_vec<-c(1:7)

Day_list<-c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")
i=1
for (k in Day_list){
  acc_data<-mcp_raw_test[day==k]
  acc_data$predicted<-as.numeric(acc_data$predicted)
  x<-error_test(acc_data$`PTF (USD/MWh)`,acc_data$predicted)
  mae_vec[i]<-x$mae
  rmse_vec[i]<-x$rmse
  i=i+1
}
mae_vec
rmse_vec
```



```{r}

mcp_raw_test$RMSE<-sqrt((mcp_raw_test$`PTF (USD/MWh)`- mcp_raw_test$predicted)^2)
mcp_raw_test$mae<-(abs(mcp_raw_test$`PTF (USD/MWh)`- mcp_raw_test$predicted))

```

