---
title: "IE 492 MCP Firts Look"
author: "Alican Yılmaz"
date: "3/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r READ, warning=FALSE, message=FALSE}
library('ggplot2')
library('forecast')
library('tseries')
library("data.table")
library("lubridate")
library("zoo")
library("plotly")
```

```{r}
setwd("/Users/Alican/Desktop")
mcp_raw = fread('MCP-01012017-29032021.csv', header=TRUE)
mcp_raw[,Date:=as.Date(Date,"%d/%m/%Y")]
```


```{r}
daily_MCP=mcp_raw[,list(daily_MCP=mean(`MCP (TL/MWh)`,na.rm=T)),by=list(Date)]
```

```{r}
plt<-ggplot(daily_MCP, aes(Date, daily_MCP)) + geom_line()  + labs(x="Date",y="MCP TL/Mwh)",title="Daily Mean mcp")

ggplotly(plt)
```

Variance not constant, there is trend for sure. Also, outliers exist. Transformation is necessary?

Currency also has an effect maybe? (2018 Currency crisis)

```{r acf/pacf}

ggAcf(daily_MCP$daily_MCP, lag=48)


```


Seasonality at weekly level.



```{r}
# Plotting a portion of daily data
ggplot(data = daily_MCP[1:28], aes(x = Date, y =daily_MCP )) +
  geom_line(color = "darkred") +
  labs(title = "Daily MCP(01.01.2017-28.01.2017) ",
       x = "Date",
       y= "MCP TL/(MWh)") +
  scale_x_date(date_minor_breaks = "1 day", date_breaks = "2 days", date_labels = "%d %b") +
  theme_minimal()
```




```{r}
plt<-ggplot(data = mcp_raw[1:96], aes(x = index(mcp_raw[1:96]), y = `MCP (TL/MWh)` )) +
  geom_line(color = "darkred") +
  labs(title = "Hourly MCP(01.01.2017-04.01.2017) ",
       x = "Date",
       y= "MCP TL/(MWh)") +
  theme_minimal()
ggplotly(plt)
```


There seems to be a seasonality at hourly level.

```{r ACF/PACF}
ggAcf(mcp_raw$`MCP (TL/MWh)`, lag=72)

```

Daily seasonality.

Weekends & holidays might affect.(But weekends might be redundant if weekly seasonality is put as input.)
Electricity Demand
Electricity Load
Temperature
Available capacity
Codes of Days
Supply & Demand of Electricity(Only shows one hour. How to download all?)
Weather conditions & Fuel Prices

```{r Data manipulation weekend and day code is added}
mcp_raw$day <- weekdays(as.Date(mcp_raw$Date))
mcp_raw$IsWeekend <- mcp_raw$day %in% c("Saturday","Sunday")
mcp_raw<-transform(mcp_raw ,IsWeekend = ifelse(mcp_raw$day %in% c("Saturday","Sunday") ,1,0 ))

```


# Web Scraping For Fuel Prices

```{r Fuel Prices in Turkey}
#install.packages("rvest")
#install.packages("dplyr")
library(rvest)
library(dplyr)
```


```{r}
link="https://www.tppd.com.tr/en/former-oil-prices?id=34&county=413&StartDate=01.01.2017&EndDate=29.03.2021"
page=read_html(link)

date<-page %>% html_nodes("td:nth-child(1)") %>% html_text()
fuel_price<-as.numeric(page %>% html_nodes("td:nth-child(6)") %>% html_text())
fuel_price_df<-data.frame(date,fuel_price,stringsAsFactors = FALSE)
fuel_price_df$date<-as.Date(fuel_price_df$date, format =  "%d %B %Y")
```

Some of the days are missing. We could interpolate them?


## ARIMA: Benchmark Purposes

Box-Jenkins Methodology will be used:

```{r}
adf.test(mcp_raw$`MCP (TL/MWh)`)
```

Data is stationary at 0.01 significance level.

```{r}
ggAcf(mcp_raw$`MCP (TL/MWh)`,lag.max = 168)
```

```{r}
ggPacf(mcp_raw$`MCP (TL/MWh)`,lag.max = 168)

```


We know that there is weekly seasonality from the daily plot.

```{r}
daily_mcp_ts<-ts(daily_MCP$daily_MCP,frequency = 7)
fitted<-auto.arima(daily_mcp_ts,seasonal=TRUE,trace = TRUE)
fitted
```


```{r}
daily_mcp_ts<-ts(daily_MCP$daily_MCP,frequency = 7)
fitted<-arima(daily_mcp_ts,order=c(3,1,1),seasonal=c(2,0,0))
fitted
```

ARIMA(3,1,1)(2,0,0)[7] I want to fit hourly version of this. # freq=24 yaz.

OR

Model each hour seperately.



## MAPE 

For one week:

Train: 2017-2020
Test: 3 Week

```{r}
# Error test
```



